---
title: "Tax convergence"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 10
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>

```{r loading_libraries, include=FALSE}

library(dplyr)
library(ggplot2)
library(gdx)
library(mip)
library(plotly)
library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```

```{r setup, echo=FALSE, include=FALSE}

# placeholder for absent/empty variables (ggplotly errors on zero-row data)
emptyPlot <- function(var, df = NULL) {
  cat(paste0("str() of raw data for '", var, "':\n"))
  str(df)
  p <- ggplot() +
    annotate("text", x = 0, y = 0, label = paste0("no data for '", var, "' in this gdx")) +
    theme_void()
  ggplotly(p)
}

booleanColor <- c("yes" = "#00BFC4", "no" = "#F8766D")

# tolerance is hardcoded to 0.001 (0.1% of GDP) in 80_optimization/nash/postsolve.gms.
# Read p80_taxConvTol from the gdx if it has been exported there, otherwise fall back.
tol <- suppressWarnings(as.numeric(gdx::readGDX(params$gdx, "p80_taxConvTol", react = "silent")))
if (length(tol) == 0 || is.na(tol)) tol <- 0.001

df <- mip::getPlotData("p80_convNashTaxrev_iter", params$gdx)
if (nrow(df) > 0) {
  df <- df %>%
    rename("value" = "p80_convNashTaxrev_iter") %>%
    mutate(
      "iteration" := as.numeric(as.character(.data$iteration)),
      "ttot" := as.numeric(as.character(.data$ttot))
    )
}

```

The tax convergence criterion requires the **net tax revenue in every region and
period to be below `r round(100 * tol, 2)`% of GDP** (`|p80_convNashTaxrev_iter| <=
`r tol``). A non-zero tax revenue means the region is still collecting/paying a net
implicit tax that should net out at the Nash equilibrium, so any region above the band
blocks convergence.

## Deviation per region and iteration

Signed tax revenue as a share of GDP. Blue = below target (paying), red = above
(collecting); saturated cells are outside the `r round(100 * tol, 2)`% band. One panel
per period. Hover for the exact value.

```{r heatmap, results = "asis"}

if (nrow(df) == 0) {
  emptyPlot("p80_convNashTaxrev_iter", df)
} else {
  cap <- 5 * tol
  hd <- df %>%
    mutate(
      "fillv" = pmax(pmin(.data$value, cap), -cap),
      "tooltip" = paste0(
        .data$all_regi, " ", .data$ttot,
        "<br>Iteration: ", .data$iteration,
        "<br>Tax revenue: ", round(100 * .data$value, 3), "% of GDP",
        "<br>", ifelse(abs(.data$value) > tol, "outside", "within"), " +-",
        round(100 * tol, 2), "% band"
      )
    )

  p <- ggplot(hd, aes_(x = ~iteration, y = ~all_regi, fill = ~fillv, text = ~tooltip)) +
    geom_tile(color = "white", linewidth = 0.1) +
    facet_wrap(~ttot) +
    scale_fill_gradient2(
      low = "#2c7bb6", mid = "#ffffbf", high = "#d7191c", midpoint = 0,
      name = "share of GDP", labels = scales::percent
    ) +
    theme_minimal() +
    labs(x = "iteration", y = NULL) +
    theme(panel.grid = element_blank())

  ggplotly(p, tooltip = c("text"))
}

```

## Convergence trajectory per region

Tax revenue over iterations, one panel per region (free y-axis). The dashed lines mark
the +-`r round(100 * tol, 2)`% tolerance band; colour distinguishes the target period.

```{r lines, results = "asis"}

if (nrow(df) == 0) {
  emptyPlot("p80_convNashTaxrev_iter", df)
} else {
  p <- ggplot(df, aes_(x = ~iteration, y = ~value, color = ~factor(ttot),
                       group = ~interaction(all_regi, ttot))) +
    geom_hline(yintercept = c(-tol, tol), linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = 0, color = "grey80") +
    geom_line(alpha = 0.7) +
    geom_point(size = 0.8, alpha = 0.7) +
    facet_wrap(~all_regi, scales = "free_y") +
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    labs(x = "iteration", y = "tax revenue (share of GDP)", color = "period")

  ggplotly(p)
}

```

## Worst offender per iteration

For each iteration, the single region/period furthest outside the tolerance band -
the cell that most needs attention.

```{r worst, results = "asis"}

if (nrow(df) == 0) {
  cat("no data")
} else {
  worst <- df %>%
    group_by(.data$iteration) %>%
    slice(which.max(abs(.data$value))) %>%
    ungroup() %>%
    mutate(
      "converged" = ifelse(abs(.data$value) > tol, "no", "yes"),
      "abs_share" = abs(.data$value)
    ) %>%
    arrange(.data$iteration)

  knitr::kable(
    worst %>%
      transmute(
        iteration = .data$iteration,
        region = .data$all_regi,
        period = .data$ttot,
        `tax revenue (% of GDP)` = round(100 * .data$value, 3),
        converged = .data$converged
      ),
    row.names = FALSE
  )
}

```
