---
title: "Convergence overview"
output:
  html_document:
    toc: false
    toc_float: true
    code_folding: hide
params:
  gdx: "fulldata.gdx"
  warning: false
  message: false
  figWidth: 12
---

<style>
  #header {
    display: none !important
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>


```{r loading_libraries, include=FALSE}

library(remind2)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)

```


## Convergence Criterias

Master overview of every convergence criterion. Each **row** is one criterion and the
**x-axis** is the Nash iteration; each **dot** shows whether that criterion was met at
that iteration (teal = met, red = not; the top feasibility row also uses amber for
skipped / feasible-but-not-optimal, and the emission-market row uses amber for a
best-achievable stop). The run converges only once **every active criterion is teal in
the same iteration**. A **black** y-axis label marks an active criterion; a **grey**
label an inactive one. Hover any dot for the underlying value and the regions/periods
still failing.

```{r results = "asis", fig.height=10}

diag <- plotNashConvergence(gdx = params$gdx)
htmltools::tagList(diag$plot)

```

## Trade summary

Detail behind the trade-surplus row above: the worst market surplus of each traded good,
per period (facet columns / rows) across iterations (x-axis). The **coloured line** is
the actual surplus and the **shaded rectangle** is the allowed tolerance band — green
while the good is inside it that iteration, red when outside. A good whose line stays
outside its band is the one holding up trade convergence.

```{r results = "asis", fig.height=10}

htmltools::tagList(diag$tradeDetailPlot)

```

